---
title: "Reeves et al 2023: Analysis Script"
subtitle: "Bayesian Models"
author: "Jonathan S. Reeves"
date: "`r Sys.Date()`"
output:  pdf_document
---

This markdown document accompanies the paper entitled, "Simulation and social network analysis provide insight into the acquisition of tool behavior in hybrid macaques". 

This script can be used to analyze the data used to generate figures 2 and 3 in the text. The data can be downloaded from the Dryad repository associated with the paper. 

# Load required packages 
```{r setup, include=FALSE}
# Packages
library(rethinking)
library(igraph)
library(dplyr)
library(knitr)
library(pander)
source("../Scripts/helper_functions.R") # Some useful helper functions.
```

# User defined variables
```{r}

N_CHAINS <- 3

```

## Formal Specification of the Model
  
```{r, Formal specification}

model_spec <-   alist(
    U ~ dbinom(1, p),
    logit(p) <- a[H] + bC[H] * C + bS[H] * S,
    a[H] ~ dnorm(0,1.5),
    bC[H] ~ dlnorm(0,.3),
    bS[H] ~ dlnorm(0,.3)
  )

```

Create a place to keep all the model results for plotting later.
```{r}

models <- list()

```

## Analysis of ABM data

### ABM: Social Learning Condition

```{r social condition}

set.seed(11) # The seed is set for consistent outcomes
nodes <- read.csv("../Data/ABM_Social_Condition.csv")

dat <- list(
  U = nodes$U, # Tool user
  C = nodes$C, # EV centrality
  S = nodes$S, # Strength to tool user
  A = nodes$A, # Age
  H = nodes$H # Phenotype
)

models$social <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = 3,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE
  )



xres <- precis(models$social, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)

xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/social_precis_link.csv")
```

### ABM: Inheritance Model



```{r}

set.seed(11)

nodes <- read.csv("../Data/ABM_Inheritance_Condition.csv")

dat <- list(
  U = nodes$U,
  C = nodes$C,
  S = nodes$S,
  A = nodes$A,
  H = nodes$H
)

models$inheritance <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = N_CHAINS,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE)



xres <- precis(models$inheritance, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)

xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/inheritance_precis_link.csv")
```

## Koram Data

### Run Model 

```{r}
nodes <- read.csv('../Data/Koram.csv')

set.seed(11)

dat <- list(
  U = nodes$U,
  C = nodes$C,
  S = nodes$S,
  A = nodes$A,
  H = nodes$H
)

models$koram_edge <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = 10,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE 
  )



xres <- precis(models$koram, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)
xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/koram_precis_link.csv")
```

### Ploting the Result

```{r, fig.height=6, fig.width=6}

names(models) <- c("Social",
                   "Inheritance",
                   "Koram")


figure_1(models)

```

## Analysis of Alternative: Resource Attraction

A new list to put the alternative hypotheses data in.

```{r}
amodels <- list()
```
### ABM N Attractors = 1, Attractor Strength = 25

```{r na1ns25 model run}

nodes <- read.csv("../Data/ABM_Asocial_na1as25_Condition.csv")

set.seed(11)

dat <- list(
  U = nodes$U,
  C = nodes$C,
  S = nodes$S,
  A = nodes$A,
  H = nodes$H
)

amodels$na1as25 <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = 10,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE 
  )


xres <- precis(amodels$na1as25, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)
xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/na1as25_precis_link.csv") 
```

### ABM N Attractors = 10, Attractor Strength = 25

```{r}
nodes <- read.csv("../Data/ABM_Asocial_na10as25_Condition.csv")

set.seed(11)

dat <- list(
  U = nodes$U,
  C = nodes$C,
  S = nodes$S,
  A = nodes$A,
  H = nodes$H
)

amodels$na10as25 <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = 10,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE # only needed on macbook air M1
  )


xres <- precis(amodels$na10as25, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)
xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/na10as25_precis_link.csv") 
```

### ABM N Attractors = 200, Attractor Strength = 25

```{r}

nodes <- read.csv("../Data/ABM_Asocial_na200as25_Condition.csv")

set.seed(11)

dat <- list(
  U = nodes$U,
  C = nodes$C,
  S = nodes$S,
  A = nodes$A,
  H = nodes$H
)

amodels$na200as25 <- ulam(
  model_spec,
  data = dat, 
  chains = N_CHAINS,
  cores = 10,
  iter = 2000,
  log_lik = TRUE, 
  cmdstan = TRUE # only needed on macbook air M1
  )


xres <- precis(amodels$na200as25, depth = 2)
xres_link <- apply(xres[,c(1:4)], 2, inv_logit)
xres_link <- cbind(xres_link, xres[c("n_eff", "Rhat4")])
xres_link$n_eff <- round(xres_link$n_eff)
xres_link$Rhat4 <- round(xres_link$Rhat4,2)
write.csv(xres_link, "../Tables/csvs/na200as25_precis_link.csv") 
```


```{r fig.height=6, fig.width=6}

names(amodels) <- c("1 Attractors",
                    "10 Attractors",
                   "200 Attractors")


figure_1(amodels)


```

